\textsf{Are frontier models superhuman chemists?}

### 2024-03-17 src/tex/ms.tex

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}

@@

% Calculation vs. no calculation
Given that some of the questions in chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.

@@

\subparagraph{Oxidation states}
The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}. These exercises focus on determining the number of isomers for specific molecules. To perform these calculations, we utilize the RDKit package \cite{bento2020open}.
\subparagraph{Electron count of atoms}
Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}. These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
\subparagraph{Number of isomers}
The questions concerning the number of isomers are sourced from the website \url{https://www.cheminfo.org/}. These exercises aim to determine the number of isomers for given molecules. The calculations for these assessments are conducted using the RDKit package \cite{bento2020open}.
\subparagraph{Point group for molecules}
Questions involve predicting the point group for molecules. The ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption/tree/main}) is utilized for data generation. Given a molecule, predict its point group.
\subparagraph{SMILES-IUPAC}
Web scrapped SMILES IUPAC pairs from the PubChem database \cite{pubchem} and then framed questions involve converting SMILES notation to IUPAC nomenclature, utilizing information from the PubChem database.

\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.

>>>>

\begin{figure}
    \centering
    \includegraphics{figures/question_diversity.pdf}
    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.}}
    \label{fig:question_diversity}
    \script{plot_question_diversity.py}
\end{figure}

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}

@@

% Calculation vs. no calculation
Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.

@@

\subparagraph{Oxidation states}
The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}. These exercises focus on determining the number of isomers for specific molecules. To perform these calculations, we utilize the RDKit package \cite{rdkit}.
\subparagraph{Electron count of atoms}
Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}. These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
\subparagraph{Number of isomers}
The questions concerning the number of isomers are sourced from the website \url{https://www.cheminfo.org/}. These exercises aim to determine the number of isomers for given molecules. The calculations for these assessments are conducted using the RDKit package \cite{rdkit}.
\subparagraph{Point group for molecules}
Questions involve predicting the point group for molecules. The ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption/tree/main}) is utilized for data generation. Given a molecule, predict its point group.
\subparagraph{SMILES-IUPAC}
Web scrapped \gls{smiles} \gls{iupac} pairs from the PubChem database \cite{pubchem} and then framed questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature, utilizing information from the PubChem database.

\paragraph{Programmatically generated questions} 
\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\cite{Irwin_2012} and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms. 
We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
